---
title:    |
          | BAN432 fall 2022 
          | First group assignment
author: "Group 11"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE)

```

```{r data}

setwd('C:/Users/steef/Documents/NHH/BAN432 Applied Textual Data Analysis for Business and Finance/Assignment Group 1/BAN432-Assignment1')
load('firm_dataset.Rdata')
rm(section.7.mda)

```

```{r libraries, message = FALSE}

library(tm)
library(tidytext)
library(stringr)
library(tibble)
library(dplyr)
library(ggplot2)
library(geomtextpath)
library(purrr)
library(tidyr)
library(PlaneGeometry) # Library to find intersections
library(scales) # Put y scale in %

## Extract code from Rmd
# knitr::purl(input = 'Group_project.Rmd',
#             output = 'Group_project.R')

```

```{r data-cleaning}

# Clean the data: remove punctuation, remove numbers, remove stopwords, make lower case, only consider
# terms with 3 to 20 letters, and delete excess white spaces.

section.1.business %>%
  # Remove the \" 
  gsub('\\\"','',.) %>% 
  # Transform to lowercase
  tolower() %>% 
  removePunctuation() %>% 
  removeNumbers() %>% 
  # Remove stopwords
  removeWords(stopwords()) %>% 
  # Remove extra spaces
  gsub(' +',' ',.) %>% 
  # Remove words with 1 to 3 characters
  gsub('\\b\\w{1,3}\\s','',.) %>% 
  # Remove words with 21 or more characters
  gsub('\\b\\w{21,}\\s','',.) -> sec.1.clean 

```

# Bi-grams

```{r tokensize-bigrams}

# Transform the data into bigrams. One easy implementation using the tokenizers package:

sec.1.clean %>%
  as_tibble()  %>%
  mutate(rown = c(1:500)) %>% # create number for future merging with raw.data
  unnest_ngrams(input = 'value', 
                output = 'bigram', 
                n = 2,
                ngram_delim = '_') -> sec.1.bigrams

```

```{r make-dtm, message = FALSE}

# Make a document term matrix only including tokens that appear in more than 5 
# but less than 100 documents.

sec.1.bigrams %>%
  group_by(rown) %>%
  select(bigram) %>%
  summarise(text = paste(bigram, collapse = ' '), .groups = 'drop') %>%
  select(text) %>%
  # To not lose the document numbers
  as.vector() %>%
  .[[1]] %>%
  VectorSource(.) %>%
  Corpus(.) %>%
  DocumentTermMatrix(control = list(bound = list(global = c(5, 100)))) -> dtm

```

```{r compute-log-lik}

# For each token (bi-grams) in the corpora compute the Log-Likelihood of being in the subset of oil firms
# and the remaining. Display the top tokens. Are these tokens associated with the oil sector? For the
# remainder of the analysis, only use the 500 tokens with the highest Log-Likelihood. Log-Likelihood was
# mentioned in one of the guest lectures and you calculate it this way:
# calculate.ll <- function(a, b, c, d) {
#   e1 <- c * (a + b) / (c + d)
#   e2 <- d * (a + b) / (c + d)
#   ll <- 2 * ((a * log(a / e1)) + (b * log(b / e2)))
#   return(ll)
# }

# a = freq. of a word in the oil corpus
a = dtm[raw.data$industry.fama.french.49 == '30 Oil', ] %>%
  as.matrix() %>%
  colSums()

# b = freq. of a word in the non-oil corpus
b = dtm[raw.data$industry.fama.french.49 != '30 Oil', ] %>%
  as.matrix() %>%
  colSums()

# c = sum of all words in the oil corpus
c = dtm[raw.data$industry.fama.french.49 == '30 Oil', ] %>%
  as.matrix() %>%
  sum()

# d = sum of all words in the non-oil corpus
d = dtm[raw.data$industry.fama.french.49 != '30 Oil', ] %>%
  as.matrix() %>%
  sum()

cbind(as.data.frame(a),b) %>%
  filter((a != 0) & # If a = 0 log is not defined
           (b != 0)) %>% # If b = 0 log is not defined
  mutate(e.1 = c * (a + b) / (c + d), 
         e.2 = d * (a + b) / (c + d)) %>%
  mutate(ll = 2 * ((a * log(a / e.1)) + (b * log(b / e.2)))) %>%
  select(ll) %>%
  arrange(desc(ll)) %>%
  head(500) -> top.500.ll

top_n(top.500.ll, n = 10) %>%
  knitr::kable(caption = 'Top 10 Log-Likelihood bigrams')

imp.bigrams <- rownames(top.500.ll) 

# Use the 500 tokens with the highest Log-Likelihood
dtm <- dtm[,imp.bigrams]

```

```{r cos-sim-peers}

CosineSimilarity <- function(A, B) {
  sum(A * B) / sqrt(sum(A ^ 2) * sum(B ^ 2))
}

d <- data.frame(matrix(ncol = 18, nrow = 482))

filter(raw.data, raw.data$industry.fama.french.49 == "30 Oil") -> Oil
filter(raw.data, raw.data$industry.fama.french.49 != "30 Oil") -> notOil

for (i in 1:18) {
  for (j in 1:482) {
    d[[j, i]] <- CosineSimilarity(dtm[raw.data$cik == Oil$cik[i], ],
                                  dtm[raw.data$cik == notOil$cik[j], ])
  }
}

# Create a list for peer firms
closest <- list()

# Select top 5
for (i in 1:18) {
  which(d[, i] %in% paste(sort(d[, i], decreasing = TRUE)[1:5])) %>%
    append(closest, .) -> closest
}

# Indexes to doc numbers
map_dbl(closest, 
        ~ which(raw.data$industry.fama.french.49 != '30 Oil')[.x]) -> 
  closest.ind

```

```{r avg-return}

# Add weights 
sort(closest.ind) %>%
  as_tibble() %>%
  group_by(value) %>%
  summarize(weight = n()) %>%
  mutate(weight = weight/length(closest)) -> portfolio.composition

# Equally weighted portfolio
raw.data %>%
  filter(industry.fama.french.49 == '30 Oil') %>%
  select(return.monthly.NY.m01 : return.monthly.NY.m12) %>%
  colMeans() %>%
  as_tibble() %>%
  rename(Oil.return = value) -> Oil.ret

# Weighted peer portfolio return
raw.data[portfolio.composition$value,] %>%
  select(return.monthly.NY.m01 : return.monthly.NY.m12) %>%
  bind_cols(portfolio.composition) %>%
  select(-value) %>%
  pivot_longer(return.monthly.NY.m01 : return.monthly.NY.m12,
               names_to = 'month', values_to = 'return') %>%
  group_by(month) %>%
  summarize(portf.ret = sum(weight * return)) %>%
  as_tibble() %>%
  bind_cols(Oil.ret) -> replicating.portf

```

```{r plot-portfolios-bigrams}

# Prepare data
replicating.portf %>%
  pivot_longer(portf.ret:Oil.return, 
               names_to = 'Portfolio', values_to = 'Return') %>%
  mutate(month = gsub('return.monthly.NY.m', '', month) %>%
           as.numeric()) -> replicating.portf.longer

# Create a dataframe for filling area
replicating.portf %>%
  mutate(
    ymax = pmax(portf.ret, Oil.return),
    ymin = pmin(portf.ret, Oil.return),
    fill = portf.ret >= Oil.return,
    month = gsub('return.monthly.NY.m', '', month) %>%
           as.numeric()
  ) -> bounds


# Find the intersection between 2 lines
# INTERSECTION 1 ---------------------------------------------------------------

# Save coordinates of 2 points on portf.ret line
portf.ret.4 <- c(bounds$month[4], bounds$portf.ret[4]) # point 1
portf.ret.5 <- c(bounds$month[5], bounds$portf.ret[5]) # point 2

# Save coordinates of 2 points on Oil.return line
Oil.return.4 <- c(bounds$month[4], bounds$Oil.return[4]) # point 1
Oil.return.5 <- c(bounds$month[5], bounds$Oil.return[5]) # point 2

# Save AD1 and AS lines as objects
portf.ret <- Line$new(portf.ret.4, portf.ret.5)
Oil.return <- Line$new(Oil.return.4, Oil.return.5)

# Find the coordinates of the intersection
coord <- intersectionLineLine(portf.ret, Oil.return)

# INTERSECTION 2 ---------------------------------------------------------------

# Save coordinates of 2 points on portf.ret line
portf.ret.7 <- c(bounds$month[7], bounds$portf.ret[7]) # point 1
portf.ret.8 <- c(bounds$month[8], bounds$portf.ret[8]) # point 2

# Save coordinates of 2 points on Oil.return line
Oil.return.7 <- c(bounds$month[7], bounds$Oil.return[7]) # point 1
Oil.return.8 <- c(bounds$month[8], bounds$Oil.return[8]) # point 2

# Save AD1 and AS lines as objects
portf.ret <- Line$new(portf.ret.7, portf.ret.8)
Oil.return <- Line$new(Oil.return.7, Oil.return.8)

# Find the coordinates of the intersection
coord2 <- intersectionLineLine(portf.ret, Oil.return)

# INTERSECTION 3 ---------------------------------------------------------------

# Save coordinates of 1 points on portf.ret line (already have august)
portf.ret.9 <- c(bounds$month[9], bounds$portf.ret[9]) 

# Save coordinates of 1 point on Oil.return line (already have august)
Oil.return.9 <- c(bounds$month[9], bounds$Oil.return[9])

# Save AD1 and AS lines as objects
portf.ret <- Line$new(portf.ret.8, portf.ret.9)
Oil.return <- Line$new(Oil.return.8, Oil.return.9)

# Find the coordinates of the intersection
coord3 <- intersectionLineLine(portf.ret, Oil.return)

# ------------------------------------------------------------------------------

bind_rows(bounds,
          tibble(month = c(coord[1],coord[1], 
                           coord2[1], coord2[1],
                           coord3[1], coord3[1]), 
                 portf.ret = c(coord[2],coord[2], 
                               coord2[2], coord2[2],
                               coord3[2], coord3[2]), 
                 Oil.return = c(coord[2],coord[2], 
                                coord2[2], coord2[2],
                                coord3[2], coord3[2]),
                 ymax = c(coord[2],coord[2], 
                          coord2[2], coord2[2],
                          coord3[2], coord3[2]), 
                 ymin = c(coord[2],coord[2], 
                          coord2[2], coord2[2],
                          coord3[2], coord3[2]), 
                 fill = c(FALSE,TRUE, 
                          TRUE, FALSE,
                          FALSE, TRUE))) %>%
  arrange(month) -> bounds

ggplot() +
  geom_ribbon(data = bounds, aes(month, ymin = ymin, ymax = ymax, fill = fill)) +
  guides(linetype = "none", fill = "none") +
  labs(x = NULL, y = NULL) +
  scale_fill_manual(values = c('#B2B2B2', '#F0F0F0')) -> area_fill

area_fill +
  geom_line(data = replicating.portf.longer, 
            aes(x = month, y = Return, color = Portfolio), 
            size = 1.1, alpha = 0.9) +
  theme_minimal() +
  theme(panel.grid = element_blank()) +
  scale_x_continuous(breaks = c(1:12), 
                     labels = month.abb) +
  labs(x = 'Month',
       title = 'Figure 1: Bi-grams replicating portfolio performance') +
  scale_color_manual(values = c('#00A9F4', '#1F40E6'),
                     labels = c('Oil portfolio', 'Replicating portfolio')) +
  scale_y_continuous(labels = scales::percent)

```




```{r evaluate-performance-bigrams}

replicating.portf %>%
  mutate(RMSE = (Oil.return - portf.ret)^2) %>%
  select(RMSE) %>%
  colSums()/12 -> RMSE.bigrams

print(paste('RMSE of bi-grams replicating portfolio is', 
            round(RMSE.bigrams,5)))

```

# Uni-grams

```{r uni-grams}
sec.1.clean %>%
  VectorSource(.) %>%
  Corpus(.) %>%
  DocumentTermMatrix(control = list(bound = list(global = c(5, 100)))) -> dtm

# a = freq. of a word in the oil corpus
a = dtm[raw.data$industry.fama.french.49 == '30 Oil', ] %>%
  as.matrix() %>%
  colSums()

# b = freq. of a word in the non-oil corpus
b = dtm[raw.data$industry.fama.french.49 != '30 Oil', ] %>%
  as.matrix() %>%
  colSums()

# c = sum of all words in the oil corpus
c = dtm[raw.data$industry.fama.french.49 == '30 Oil', ] %>%
  as.matrix() %>%
  sum()

# d = sum of all words in the non-oil corpus
d = dtm[raw.data$industry.fama.french.49 != '30 Oil', ] %>%
  as.matrix() %>%
  sum()

cbind(as.data.frame(a),b) %>%
  filter((a != 0) & # If a = 0 log is not defined
           (b != 0)) %>% # If b = 0 log is not defined
  mutate(e.1 = c * (a + b) / (c + d), 
         e.2 = d * (a + b) / (c + d)) %>%
  mutate(ll = 2 * ((a * log(a / e.1)) + (b * log(b / e.2)))) %>%
  select(ll) %>%
  arrange(desc(ll)) %>%
  head(500) -> top.500.ll

top_n(top.500.ll, n = 10) %>%
  knitr::kable(caption = 'Top 10 Log-Likelihood unigrams')

imp.bigrams <- rownames(top.500.ll) 

# Use the 500 tokens with the highest Log-Likelihood
dtm <- dtm[,imp.bigrams]

CosineSimilarity <- function(A, B) {
  sum(A * B) / sqrt(sum(A ^ 2) * sum(B ^ 2))
}

d <- data.frame(matrix(ncol = 18, nrow = 482))

filter(raw.data, raw.data$industry.fama.french.49 == "30 Oil") -> Oil
filter(raw.data, raw.data$industry.fama.french.49 != "30 Oil") -> notOil

for (i in 1:18) {
  for (j in 1:482) {
    d[[j, i]] <- CosineSimilarity(dtm[raw.data$cik == Oil$cik[i], ],
                                  dtm[raw.data$cik == notOil$cik[j], ])
  }
}

# Create a list for peer firms
closest <- list()

# Select top 5
for (i in 1:18) {
  which(d[, i] %in% paste(sort(d[, i], decreasing = TRUE)[1:5])) %>%
    append(closest, .) -> closest
}

# Indexes to doc numbers
map_dbl(closest, 
        ~ which(raw.data$industry.fama.french.49 != '30 Oil')[.x]) -> 
  closest.ind

# Add weights 
sort(closest.ind) %>%
  as_tibble() %>%
  group_by(value) %>%
  summarize(weight = n()) %>%
  mutate(weight = weight/length(closest)) -> portfolio.composition

# Equally weighted portfolio
raw.data %>%
  filter(industry.fama.french.49 == '30 Oil') %>%
  select(return.monthly.NY.m01 : return.monthly.NY.m12) %>%
  colMeans() %>%
  as_tibble() %>%
  rename(Oil.return = value) -> Oil.ret

# Weighted peer portfolio return
raw.data[portfolio.composition$value,] %>%
  select(return.monthly.NY.m01 : return.monthly.NY.m12) %>%
  bind_cols(portfolio.composition) %>%
  select(-value) %>%
  pivot_longer(return.monthly.NY.m01 : return.monthly.NY.m12,
               names_to = 'month', values_to = 'return') %>%
  group_by(month) %>%
  summarize(portf.ret = sum(weight * return)) %>%
  as_tibble() %>%
  bind_cols(Oil.ret) -> replicating.portf
```



```{r plot-portfolios-unigrams}

# Prepare data
replicating.portf %>%
  pivot_longer(portf.ret:Oil.return, 
               names_to = 'Portfolio', values_to = 'Return') %>%
  mutate(month = gsub('return.monthly.NY.m', '', month) %>%
           as.numeric()) -> replicating.portf.longer

# Create a dataframe for filling area
replicating.portf %>%
  mutate(
    ymax = pmax(portf.ret, Oil.return),
    ymin = pmin(portf.ret, Oil.return),
    fill = portf.ret >= Oil.return,
    month = gsub('return.monthly.NY.m', '', month) %>%
           as.numeric()
  ) -> bounds


# Find the intersection between 2 lines
# INTERSECTION 1 ---------------------------------------------------------------

# Save coordinates of 2 points on portf.ret line
portf.ret.2 <- c(bounds$month[2], bounds$portf.ret[2]) # point 1
portf.ret.3 <- c(bounds$month[3], bounds$portf.ret[3]) # point 2

# Save coordinates of 2 points on Oil.return line
Oil.return.2 <- c(bounds$month[2], bounds$Oil.return[2]) # point 1
Oil.return.3 <- c(bounds$month[3], bounds$Oil.return[3]) # point 2

# Save AD1 and AS lines as objects
portf.ret <- Line$new(portf.ret.2, portf.ret.3)
Oil.return <- Line$new(Oil.return.2, Oil.return.3)

# Find the coordinates of the intersection
coord <- intersectionLineLine(portf.ret, Oil.return)

# INTERSECTION 2 ---------------------------------------------------------------

# Save coordinates of 2 points on portf.ret line
portf.ret.4 <- c(bounds$month[4], bounds$portf.ret[4]) # point 1
portf.ret.5 <- c(bounds$month[5], bounds$portf.ret[5]) # point 2

# Save coordinates of 2 points on Oil.return line
Oil.return.4 <- c(bounds$month[4], bounds$Oil.return[4]) # point 1
Oil.return.5 <- c(bounds$month[5], bounds$Oil.return[5]) # point 2

# Save AD1 and AS lines as objects
portf.ret <- Line$new(portf.ret.4, portf.ret.5)
Oil.return <- Line$new(Oil.return.4, Oil.return.5)

# Find the coordinates of the intersection
coord2 <- intersectionLineLine(portf.ret, Oil.return)

# INTERSECTION 3 ---------------------------------------------------------------

# Save coordinates of 1 points on portf.ret line (already have may)
portf.ret.6 <- c(bounds$month[6], bounds$portf.ret[6])

# Save coordinates of 1 point on Oil.return line (already have may)
Oil.return.6 <- c(bounds$month[6], bounds$Oil.return[6])

# Save AD1 and AS lines as objects
portf.ret <- Line$new(portf.ret.5, portf.ret.6)
Oil.return <- Line$new(Oil.return.5, Oil.return.6)

# Find the coordinates of the intersection
coord3 <- intersectionLineLine(portf.ret, Oil.return)

# INTERSECTION 4 ---------------------------------------------------------------

# Save coordinates of 1 points on portf.ret line (already have june)
portf.ret.7 <- c(bounds$month[7], bounds$portf.ret[7])

# Save coordinates of 1 point on Oil.return line (already have june)
Oil.return.7 <- c(bounds$month[7], bounds$Oil.return[7])

# Save AD1 and AS lines as objects
portf.ret <- Line$new(portf.ret.6, portf.ret.7)
Oil.return <- Line$new(Oil.return.6, Oil.return.7)

# Find the coordinates of the intersection
coord4 <- intersectionLineLine(portf.ret, Oil.return)

# ------------------------------------------------------------------------------
 
bind_rows(bounds,
          tibble(month = c(coord[1],coord[1],
                           coord2[1], coord2[1],
                           coord3[1], coord3[1],
                           coord4[1], coord4[1]
                           ),
                 portf.ret = c(coord[2],coord[2],
                               coord2[2], coord2[2],
                               coord3[2], coord3[2],
                               coord4[2], coord4[2]
                               ),
                 Oil.return = c(coord[2],coord[2],
                                coord2[2], coord2[2],
                                coord3[2], coord3[2],
                                coord4[2], coord4[2]
                                ),
                 ymax = c(coord[2],coord[2],
                          coord2[2], coord2[2],
                          coord3[2], coord3[2],
                          coord4[2], coord4[2]
                          ),
                 ymin = c(coord[2],coord[2],
                          coord2[2], coord2[2],
                          coord3[2], coord3[2],
                          coord4[2], coord4[2]
                          ),
                 fill = c(TRUE,FALSE,
                          TRUE, FALSE,
                          FALSE, TRUE,
                          TRUE, FALSE
                          ))) %>%
  arrange(month) -> bounds

ggplot() +
  geom_ribbon(data = bounds, aes(month, ymin = ymin, ymax = ymax, fill = fill)) +
  guides(linetype = "none", fill = "none") +
  labs(x = NULL, y = NULL) +
  scale_fill_manual(values = c('#B2B2B2', '#F0F0F0')) -> area_fill

area_fill +
  geom_line(data = replicating.portf.longer, 
            aes(x = month, y = Return, color = Portfolio), 
            size = 1.1, alpha = 0.9) +
  theme_minimal() +
  theme(panel.grid = element_blank()) +
  scale_x_continuous(breaks = c(1:12), 
                     labels = month.abb) +
  labs(x = 'Month',
       title = 'Figure 2: Uni-grams replicating portfolio performance') +
  scale_color_manual(values = c('#00A9F4', '#1F40E6'),
                     labels = c('Oil portfolio', 'Replicating portfolio')) +
  scale_y_continuous(labels = scales::percent)

```



```{r evaluate-unigram-performance}

replicating.portf %>%
  mutate(RMSE = (Oil.return - portf.ret)^2) %>%
  select(RMSE) %>%
  colSums()/12 -> RMSE.uni.grams

print(paste('RMSE of uni-grams replicating portfolio is', 
            round(RMSE.uni.grams,5)))

```

Oil firm tracking portfolio constructed based on the uni-grams document term matrix performs significantly better than the one constructed with the use of bi-grams, as demonstrated by the Root Mean Squared Error measure. However, the use of uni-grams or bi-grams does not affect the overall relative shape of returns as shown by the charts above.

# Optional

```{r optional}

size <- 10000

replicating.portf.rand <- data.frame(matrix(ncol = size, nrow = 12))

for (i in 1:size){
  # Sample 38 stocks
  sample(which(raw.data$industry.fama.french.49 != '30 Oil'),
         nrow(portfolio.composition), replace = TRUE) %>% # Allow twice

  as_tibble() %>%
  group_by(value) %>%
  summarize(weight = n()) %>%
  mutate(weight = weight/length(closest)) -> portfolio.composition

  raw.data[portfolio.composition$value,] %>%
  select(return.monthly.NY.m01 : return.monthly.NY.m12) %>%
  bind_cols(portfolio.composition) %>%
  select(-value) %>%
  pivot_longer(return.monthly.NY.m01 : return.monthly.NY.m12,
               names_to = 'month', values_to = 'return') %>%
  group_by(month) %>%
  summarize(portf.ret = sum(weight * return)) %>%
  as_tibble() %>%
    select(portf.ret) -> replicating.portf.rand[,i]
}

(replicating.portf.rand - replicating.portf$Oil.return)^2 %>%
  colSums() / 12 -> RMSE.rand

RMSE.rand %>%
  as_tibble() %>%
  ggplot(aes(value)) +
  geom_histogram(bins = 30, fill = '#00A9F4', alpha = 0.5) +
  geom_textvline(label = 'Bigrams', 
                 xintercept = RMSE.bigrams, vjust = 1.3,
                 color = '#B2B2B2') +
  geom_textvline(label = 'Uni-grams', 
                 xintercept = RMSE.uni.grams, vjust = 1.3,
                 color = '#B2B2B2') +
  geom_textvline(label = 'Minimum RMSE random portfolio', 
                 xintercept = min(RMSE.rand), vjust = 1.3,
                 color = '#1F40E6') +
  theme_minimal() +
  theme(panel.grid = element_blank()) +
  labs(title = 'Figure 3: Random portfolios RMSE distribution')

```
